---
title: "Case Study - Bank Marketing"
output: html_notebook
---

Goal: predict if the client will subscribe a term deposit (variable y)
  Note: Y will be a binomial result, either yes or no

Step 1) Import Data
```{r}
getwd()
setwd( paste(getwd() , "/BankAnalysis_Study"))
library(readr)
library(dplyr)
library(ggplot2)
 bank <- bank <- read_delim("bank/bank.csv", 
    ";", escape_double = FALSE, 
    col_types = cols(age = col_double(), 
         balance = col_double(), 
         campaign = col_integer(), 
         contact = col_factor(levels = c("cellular", "telephone")), 
         day = col_integer(), 
         default = col_factor(levels = c("yes","no")), 
         duration = col_double(), 
         education = col_factor(levels = c("primary" , "secondary" , "tertiary")),
        # education = col_factor(levels = c("basic.4y",  "basic.6y", "basic.9y", "high.school", 
        #     "illiterate", "professional.course", 
        #     "university.degree")),
         housing = col_factor(levels = c("yes", "no")),
         job = col_factor(levels = c("admin.", 
             "blue-collar", "entrepreneur", 
             "housemaid", "management", 
             "retired", "self-employed", 
             "services", "student", "technician", 
             "unemployed")), 
         loan = col_factor(levels = c("yes", "no")), 
         marital = col_factor(levels = c("divorced", "married", "single")), 
         month = col_factor(levels = c("jan", 
             "feb", "mar", "apr", "may", "jun", 
             "jul", "aug", "sep", "oct", "nov", 
             "dec")), 
         pdays = col_number(), 
         poutcome = col_factor(levels = c("failure", "nonexistent", "success" , "other")), 
         previous = col_number(), 
         y = col_factor(levels = c("yes","no"))), 
    na = "unknown",
    trim_ws = TRUE)
 
bank_full <- read_delim("bank/bank-full.csv", 
     ";", 
     escape_double = FALSE, 
     col_types = cols(age = col_integer(), 
         balance = col_double(), campaign = col_integer(), 
         contact = col_character(), day = col_integer(), 
         default = col_factor(levels = c("yes", 
             "no")), duration = col_double(), 
         education = col_character(), housing = col_factor(levels = c("yes", 
             "no")), job = col_character(), 
         marital = col_character(), month = col_character(), 
         pdays = col_double(), poutcome = col_character(), 
         y = col_factor(levels = c("yes", 
             "no"))), 
     na = "unknown", 
     trim_ws = TRUE)

banka <- read_delim("bank-additional/bank-additional/bank-additional.csv", 
    ";", escape_double = FALSE, 
    col_types = cols(age = col_double(), 
          job = col_factor(levels = c("admin.", 
             "blue-collar", "entrepreneur", 
             "housemaid", "management", 
             "retired", "self-employed", 
             "services", "student", "technician", 
             "unemployed" , "unknown")), 
        marital = col_factor(levels = c("divorced", "married", "single" , "unknown")), 
        education = col_factor(levels = c("basic.4y",  "basic.6y", "basic.9y", "high.school", 
             "illiterate", "professional.course", 
             "university.degree" , "unknown")),
        default = col_factor(levels = c("yes","no", "unknown")), 
         housing = col_factor(levels = c("yes", "no" , "unknown")),
         loan = col_factor(levels = c("yes", "no", "unknown" )), 
         contact = col_factor(levels = c("cellular", "telephone")), 
         month = col_factor(levels = c("jan", 
             "feb", "mar", "apr", "may", "jun", 
             "jul", "aug", "sep", "oct", "nov", 
             "dec")),
        day_of_week = col_factor(levels = c("mon" , "tue","wed","thu", "fri")),
        duration = col_double(), 
        campaign = col_integer(), 
        pdays = col_integer(), #NOTE: 999 is a null value
        previous = col_integer(),
        poutcome = col_factor(levels = c("failure", "nonexistent", "success")), 
        emp.var.rate = col_double(),
        cons.price.idx = col_double(),
        cons.conf.idx = col_double(),
        euribor3m = col_double(),
        y = col_factor(levels = c("yes","no"))), 
    #na = "unknown",
    trim_ws = TRUE)
 
banka_full <- read_delim("bank-additional/bank-additional/bank-additional-full.csv", 
     ";", 
     escape_double = FALSE, 
     col_types = cols(age = col_double(), 
          job = col_factor(levels = c("admin.", 
             "blue-collar", "entrepreneur", 
             "housemaid", "management", 
             "retired", "self-employed", 
             "services", "student", "technician", 
             "unemployed" , "unknown")), 
        marital = col_factor(levels = c("divorced", "married", "single" , "unknown")), 
        education = col_factor(levels = c("basic.4y",  "basic.6y", "basic.9y", "high.school", 
             "illiterate", "professional.course", 
             "university.degree" , "unknown")),
        default = col_factor(levels = c("yes","no", "unknown")), 
         housing = col_factor(levels = c("yes", "no" , "unknown")),
         loan = col_factor(levels = c("yes", "no", "unknown" )), 
         contact = col_factor(levels = c("cellular", "telephone")), 
         month = col_factor(levels = c("jan", 
             "feb", "mar", "apr", "may", "jun", 
             "jul", "aug", "sep", "oct", "nov", 
             "dec")),
        day_of_week = col_factor(levels = c("mon" , "tue","wed","thu", "fri")),
        duration = col_double(), 
        campaign = col_integer(), 
        pdays = col_integer(), #NOTE: 999 is a null value
        previous = col_integer(),
        poutcome = col_factor(levels = c("failure", "nonexistent", "success")), 
        emp.var.rate = col_double(),
        cons.price.idx = col_double(),
        cons.conf.idx = col_double(),
        euribor3m = col_double(),
        y = col_factor(levels = c("yes","no"))),  
    #na = "unknown",
    trim_ws = TRUE)


```
Add unqiue ID :

```{r}
banka$ID <- seq.int(nrow(banka))
banka_full$ID <- seq.int(nrow(banka_full))
```




Deal with NULL/missing values
```{r}
summary(banka$pdays) 
banka$pdays[banka$pdays == 999] <- NA

summary(banka_full$pdays) 
banka_full$pdays[banka_full$pdays == 999] <- NA

na_count <-sapply(banka, function(y) sum(length(which(is.na(y)))))
na_count 


##too many pday values are 999 or NA, create a cateogrical column instead
banka$contactedprev <- 'Yes'
banka$contactedprev[ banka$pdays == 999 ] <- 'No'
banka$contactedprev <-factor(banka$contactedprev)
summary(banka$contactedprev)

banka_full$contactedprev <- 'Yes'
banka_full$contactedprev[ banka_full$pdays == 999 ] <- 'No' 
banka_full$contactedprev <-factor(banka_full$contactedprev)
summary(banka_full$contactedprev)

#change y column to be 1's and 0's to help model?
banka$output <- 0 
#banka$output[banka$y == "no"] <- 0
banka$output[banka$y == "yes"] <- 1

#change y column to be 1's and 0's to help model?
banka_full$output <- 0 
banka_full$output[banka$y == "yes"] <- 1
```
Create first model using binomial glm. 
```{r}

glm.all <- glm( output ~ age + 
                  job +
                  marital +
                  education +
                   default +
                   housing +
                   loan +
                   contact +
                   month +
                   day_of_week +
                   #duration +
                   campaign +
                   previous +
                   poutcome +
                  # contactedprev +
                   emp.var.rate +
                   cons.price.idx +
                   cons.conf.idx +
                   nr.employed
                , data = banka  , family = binomial) #Note: Used binomial because output is binary

summary(glm.all)
```

Predict. Store predicted in results data set , and compare to actual
```{r}
#predict if the client subsribed to a term deposit
glm.probs <- predict(glm.all, banka_full, type = "response")
#getting "Prediction from a rank-deficient fit may be misleading" message, because of so many input variables
SubscribedPred =rep( 0 , nrow(banka_full))
SubscribedPred[ glm.probs > .5] = 1
results <- data.frame( banka_full$ID, SubscribedPred, banka_full$output)
results <- results %>% mutate( match = ifelse( SubscribedPred == banka_full.output , 1 , 0))
head(results)

#how well does glm.probs fit test
table(results$SubscribedPred, banka_full$output)
sum(results$match)/nrow(results)
#86% correct when predictions for bank-additional-full are compared to actual
```

Now, create a model with only the significant variables:

```{r}

#try to better model by switching what variables are used in model
glm.new <- glm( output ~ 
                  #age + 
                  job +
                  #marital +
                  #education +
                  #default +
                  #housing +
                  #loan +
                  contact +
                  month +
                  day_of_week +
                  poutcome +
                  #duration +
                  #campaign +
                  #previous +
                  #contactedprev +
                  emp.var.rate +
                  cons.price.idx +
                  cons.conf.idx 
                  #nr.employed
                , data = banka , family = binomial) #Note: Used binomial because output is binary



```
New model:
```{r}
summary(glm.new)
```
Things to note: Less likely to subscribe to a term deposit on a monday. Also less likely to subscribe Jan-Mar vs any other month. More likely to subscribe if previously contacted. Blue collar are the mosth likely job to say yes to subscribe.
```{r}
#predict if the client subsribed to a term deposit
glm.probs <- predict(glm.new, banka_full, type = "response")
SubscribedPred =rep( 0 , nrow(banka_full))
SubscribedPred[ glm.probs > .5] = 1
results <- data.frame( banka_full$ID, SubscribedPred, banka_full$output)
results <- results %>% mutate( match = ifelse( SubscribedPred == banka_full.output , 1 , 0))
head(results)

#how well does glm.probs fit test
table(results$SubscribedPred, banka_full$output)
sum(results$match)/nrow(results)
# 86% correct when predictions for bank-additional-full are compared to actual
```

