---
title: "Case Study - Bank Marketing"
output: html_notebook
---

Goal: predict if the client will subscribe a term deposit (variable y)
  Note: Y will be a binomial result, either yes or no

MP , 5/27/2020
--

Import and factor Data:
```{r}
getwd()
setwd( paste(getwd() , "/BankAnalysis_Study/bank-additional/bank-additional/"))
library(readr)
library(dplyr)
library(ggplot2)

#------------------------------------------------------------------------------------------
banka <- read_delim("bank-additional/bank-additional/bank-additional.csv", 
    ";", escape_double = FALSE, 
    col_types = cols(age = col_double(), 
          job = col_factor(levels = c("admin.", 
             "blue-collar", "entrepreneur", 
             "housemaid", "management", 
             "retired", "self-employed", 
             "services", "student", "technician", 
             "unemployed" , "unknown")), 
        marital = col_factor(levels = c("divorced", "married", "single" , "unknown")), 
        education = col_factor(levels = c("basic.4y",  "basic.6y", "basic.9y", "high.school", 
             "illiterate", "professional.course", 
             "university.degree" , "unknown")),
        default = col_factor(levels = c("yes","no", "unknown")), 
         housing = col_factor(levels = c("yes", "no" , "unknown")),
         loan = col_factor(levels = c("yes", "no", "unknown" )), 
         contact = col_factor(levels = c("cellular", "telephone")), 
         month = col_factor(levels = c("jan", 
             "feb", "mar", "apr", "may", "jun", 
             "jul", "aug", "sep", "oct", "nov", 
             "dec")),
        day_of_week = col_factor(levels = c("mon" , "tue","wed","thu", "fri")),
        duration = col_double(), 
        campaign = col_integer(), 
        pdays = col_integer(), #NOTE: 999 is a null value
        previous = col_integer(),
        poutcome = col_factor(levels = c("failure", "nonexistent", "success")), 
        emp.var.rate = col_double(),
        cons.price.idx = col_double(),
        cons.conf.idx = col_double(),
        euribor3m = col_double(),
        y = col_factor(levels = c("yes","no"))), 
    #na = "unknown",
    trim_ws = TRUE)
 
banka_full <- read_delim("bank-additional/bank-additional/bank-additional-full.csv", 
     ";", 
     escape_double = FALSE, 
     col_types = cols(age = col_double(), 
          job = col_factor(levels = c("admin.", 
             "blue-collar", "entrepreneur", 
             "housemaid", "management", 
             "retired", "self-employed", 
             "services", "student", "technician", 
             "unemployed" , "unknown")), 
        marital = col_factor(levels = c("divorced", "married", "single" , "unknown")), 
        education = col_factor(levels = c("basic.4y",  "basic.6y", "basic.9y", "high.school", 
             "illiterate", "professional.course", 
             "university.degree" , "unknown")),
        default = col_factor(levels = c("yes","no", "unknown")), 
         housing = col_factor(levels = c("yes", "no" , "unknown")),
         loan = col_factor(levels = c("yes", "no", "unknown" )), 
         contact = col_factor(levels = c("cellular", "telephone")), 
         month = col_factor(levels = c("jan", 
             "feb", "mar", "apr", "may", "jun", 
             "jul", "aug", "sep", "oct", "nov", 
             "dec")),
        day_of_week = col_factor(levels = c("mon" , "tue","wed","thu", "fri")),
        duration = col_double(), 
        campaign = col_integer(), 
        pdays = col_integer(), #NOTE: 999 is a null value
        previous = col_integer(),
        poutcome = col_factor(levels = c("failure", "nonexistent", "success")), 
        emp.var.rate = col_double(),
        cons.price.idx = col_double(),
        cons.conf.idx = col_double(),
        euribor3m = col_double(),
        y = col_factor(levels = c("yes","no"))),  
    #na = "unknown",
    trim_ws = TRUE)

summary(banka)
summary(banka_full)
```

Add unqiue ID to each dataset:
```{r}
banka$ID <- seq.int(nrow(banka))
banka_full$ID <- seq.int(nrow(banka_full))
```


Look at null/Missing values:
```{r}


#The below code changes 999 to be NA's. Decided this is unneccesary and commented this out because I'd rather make a categorical column to represent pdays. Too many NA's arn't good for , and leaving the 999 would skew that data if it was treaded as numerical.
# summary(banka$pdays) 
# banka$pdays[banka$pdays == 999] <- NA
# 
# summary(banka_full$pdays) 
# banka_full$pdays[banka_full$pdays == 999] <- NA
# 
# na_count <-sapply(banka, function(y) sum(length(which(is.na(y)))))
# na_count 


##too many pday values are 999 or NA, creating a cateogrical column instead
banka$contactedprev <- 'Yes'
banka$contactedprev[ banka$pdays == 999 ] <- 'No'
banka$contactedprev <-factor(banka$contactedprev)
summary(banka$contactedprev)

banka_full$contactedprev <- 'Yes'
banka_full$contactedprev[ banka_full$pdays == 999 ] <- 'No' 
banka_full$contactedprev <-factor(banka_full$contactedprev)
summary(banka_full$contactedprev)

```

```{r}
#change y column to be 1's and 0's for binomial model
banka$output <- 0 
banka$output[banka$y == "yes"] <- 1

#change y column to be 1's and 0's for the model
banka_full$output <- 0 
banka_full$output[banka$y == "yes"] <- 1

```
Going to treat 'banka' as a train dataset , banka_full as a test dataset. 'banka' will be used to train the model, and the model will then be tested on banka_full, to gauge how accurate it is.


Create first model using binomial glm (general linear model) to look at significance of variables :
```{r}

glm.all <- glm( output ~ age + 
                  job +
                  marital +
                  education +
                   default +
                   housing +
                   loan +
                   contact +
                   month +
                   day_of_week +
                   #duration +
                   campaign +
                   previous +
                   poutcome +
                   contactedprev +
                   emp.var.rate +
                   cons.price.idx +
                   cons.conf.idx +
                   nr.employed
                , data = banka  , family = binomial) #Note: Used binomial because result is binary 

summary(glm.all)
```
Things to note: 
-Less likely to subscribe to a term deposit on a monday. 
-Highly significant if contacted on a telephone number vs cellphone
-Management and self-employed seem to be more significant than other occupations
-Social/Economic variables are significant

 

Now, create a model with only the significant variables:

```{r}

#try to better model by switching what variables are used in model. Only use top 5 significant values.
glm.new <- glm( output ~ 
                  #age + 
                  #job +
                  #marital +
                  #education +
                  #default +
                  #housing +
                  #loan +
                  contact +
                  month +
                  #day_of_week +
                  #poutcome +
                  #duration +
                  #campaign 
                  #previous +
                  #contactedprev +
                  emp.var.rate +
                  cons.price.idx +
                  cons.conf.idx 
                  #nr.employed
                , data = banka , family = binomial) #Note: Used binomial because output is binary



```
New model:
```{r}
summary(glm.new)
```

Predict. Store predicted in results data set , and compare to actual:
```{r}
#predict if the client subsribed to a term deposit
glm.probs <- predict(glm.new, banka_full, type = "response")
SubscribedPred =rep( 0 , nrow(banka_full))
SubscribedPred[ glm.probs > .5] = 1
results <- data.frame( banka_full$ID, SubscribedPred, banka_full$output)
results <- results %>% mutate( match = ifelse( SubscribedPred == banka_full.output , 1 , 0))
head(results)

#how well does glm.probs fit test
table(results$SubscribedPred, banka_full$output)
sum(results$match)/nrow(results)
# 86.89% correct when predictions for bank-additional-full are compared to actual
```


Model "glm.new" has an acceptable result of 86.89% accuracy. This number, shows that the model fits the dataset, but is not too close to overfitting.


Plot relationship between clients subscribing to a term deposit(y) and the metrics used in the model:
```{r}

#prelimary graph, to draw initial conclusions
plot( banka_full$y ~  banka_full$poutcome + banka_full$month + banka_full$contact +  
        banka_full$campaign + banka_full$job)

#(more colorful) graphs to show relationships between certain variables
library(ggplot2)
library(gridExtra)
library(scales)
#ggplot(banka_full) + 
  #geom_col(aes(x=month   fill = y))
  #geom_bar( aes(fill=y))
  #stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE)

#crate graphs for differenct input variables
#using bar plot for categorical varibles
age <- ggplot(banka, aes(fill=y , x = age)) +
  geom_bar(position = "stack" , stat= "count")  +
   geom_text( aes(label = stat(count)) , stat = 'count', size = 2.5, position = position_stack(vjust = 0.5))

poutcome <- ggplot(banka, aes(fill=y , x = day_of_week)) +
  geom_bar(position = "stack" , stat= "count")  +
   geom_text( aes(label = stat(count)) , stat = 'count', size = 2.5, position = position_stack(vjust = 0.5))

month <- ggplot(banka, aes(fill=y , x = month)) +
  geom_bar(position = "stack" , stat= "count")  +
   geom_text( aes(label = stat(count)) , stat = 'count', size = 2.5, position = position_stack(vjust = 0.5))

poutcome <- ggplot(banka, aes(fill=y , x = poutcome)) +
  geom_bar(position = "stack" , stat= "count")  +
   geom_text( aes(label = stat(count)) , stat = 'count', size = 2.5, position = position_stack(vjust = 0.5))

contact <- ggplot(banka, aes(fill=y , x = contact)) +
  geom_bar(position = "stack" , stat= "count")  +
   geom_text( aes(label = stat(count)) , stat = 'count', size = 2.5, position = position_stack(vjust = 0.5))

campaign <- ggplot(banka, aes(fill=y , x = campaign)) +
  geom_bar(position = "stack" , stat= "count")  +
   geom_text( aes(label = stat(count)) , stat = 'count', size = 2.5, position = position_stack(vjust = 0.5))

job <- ggplot(banka, aes(fill=y , x = job )) +
  geom_bar(position = "stack" , stat= "count" ) +
  theme(axis.text.x = element_text(angle = 90))  +
   geom_text( aes(label = stat(count)) , stat = 'count', size = 2.5, position = position_stack(vjust = 0.5))

#using line instead of bar chart for numerical variables
cons.conf.idx <- ggplot(banka, aes( x = cons.conf.idx)) +
  geom_line(  aes(color = y),  stat = "count") +
  geom_point( aes(color = y), stat = "count")
  

cons.price.idx <- ggplot(banka, aes( x = cons.price.idx)) +
  geom_line(  aes(color = y),  stat = "count") +
  geom_point( aes(color = y), stat = "count")

  
emp.var.rate <- ggplot(banka, aes( x = emp.var.rate)) +
  geom_line(  aes(color = y),  stat = "count") +
  geom_point( aes(color = y), stat = "count") 

 
 #graphs for significant variables  
month
contact
cons.conf.idx
cons.price.idx
emp.var.rate





```



